---
title: "Simulations Pre-Class Project"
date: "Due October 24, 2018 at 10:00pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:


With this project we will simulate a famoues probability problem. This will not require knowledge of probability or statistics but only the logic to follow the steps in order to simulate this problem. This is one way to solve problems by using the computer. 

 1. **Gambler's Ruin**: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:
    a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet. 
   
```{r}

    bankroll = 1000
    bet = 100
    tracker <- c()
fair_game()
  tracker
  
  #fair_game is the function that calculates whether after 100 bets, you have "busted"
  fair_game <- function(bankroll=1000, bet=100, n = 100) {
     for(i in 1:n){
        bankroll <- (bankroll - bet)
        if(bankroll <= 0){
          tracker <- rbind(tracker, ("Busted"))
          return(tracker)
          stop()
        } else if(i == n & bankroll >0){
             tracker <- rbind(tracker, ("Won"))
        } else if(rbinom(1, 1, .5) == 1) {
          bankroll <- bankroll + (2 * bet)
        } else if(rbinom(1, 1, .5) == 1){
            bankroll <- bankroll          
        } 
     }
   return(tracker)}
  
#check is our simulation of 5000 iterations of the game.  
check <- c()   
check <- replicate(5000, fair_game())
check

#this for loop will count the # of times "busted" in simulation
count <- 0
for(i in 1:length(check)){
  if(check[i] == "Busted"){
    count = count+1
  }
}
count
#Probability busted: (about 37%)
count/length(check)

```    
    b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly. 
    
```{r}
check2 <- replicate(5000, fair_game(bankroll = 1000, bet = 100, n = 500))

#this for loop will count the # of times "busted" in simulation
count <- 0
for(i in 1:length(check2)){
  if(check2[i] == "Busted"){
    count = count+1
  }
}

count
#Probability busted: (about 68%)
count/length(check)

```
    c. the mean time you go bust, given that you go bust within the first 5000 hands.
    
```{r}

meanTimeBust <- function(bankroll=1000, bet=100, n = 5000) {
     for(i in 1:n){
        bankroll <- (bankroll - bet)
        if(bankroll <= 0){
          tracker <- i
          return(tracker)
          stop()
        } else if(i == n & bankroll >0){
             tracker <- ("Won")
        } else if(rbinom(1, 1, .5) == 1) {
          bankroll <- bankroll + (2 * bet)
        } else if(rbinom(1, 1, .5) == 1){
            bankroll <- bankroll          
        } 
     }
   return(tracker)}

findMean<-replicate(5000, meanTimeBust())

#keep only non-"Won" values
bustedMean <- c()
for(i in 1:length(findMean)){
  if(findMean[i] != "Won"){
    bustedMean[length(bustedMean)+1] <- findMean[i]
  }
}

#find mean (467 when I ran it)
mean(as.numeric(bustedMean))



```
    d. the mean and variance of your bankroll after 100 hands (including busts).
    
```{r}

timeBust <- function(bankroll=1000, bet=100, n = 100) {
     for(i in 1:n){
        bankroll <- (bankroll - bet)
        if(bankroll <= 0){
          tracker <- 0
          return(tracker)
          stop()
        } else if(i == n & bankroll >0){
             tracker <- bankroll
        } else if(rbinom(1, 1, .5) == 1) {
          bankroll <- bankroll + (2 * bet)
        } else if(rbinom(1, 1, .5) == 1){
            bankroll <- bankroll          
        } 
     }
   return(tracker)}

findMeanVar<-replicate(5000, timeBust())
findMeanVar

#find mean and variance of bankroll
meanBankroll <- mean(findMeanVar)
varBankroll <- var(findMeanVar)

#Mean and Variance of bankroll printed 
meanBankroll
varBankroll
```
    e. the mean and variance of your bankroll after 500 hands (including busts).
    
```{r}

findMeanVar500Hands <- replicate(5000, timeBust(bankroll=1000, bet=100, n = 500))

#find mean and varaince of bankroll
meanBankRoll500 <- mean(findMeanVar500Hands)
varBankRoll500 <- var(findMeanVar500Hands)

#Mean and Variance Printed
meanBankRoll500
varBankRoll500

```
 
Note: you *must* stop playing if your player has gone bust. How will you handle this in the `for` loop?

2. **Markov Chains**. Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100.
    a. Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation.
    markov_gamble <- function(x=0.48, y){
    if()
    }
    
    
    
    b. Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically?
    c. Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it. 
