# pre-class


Make sure you commit this often with meaningful messages. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
# naming the f1 function checkPrefix because it takes a string and prefix, and based on the number of characters in prefix value, determines if the prefix matches the prefix of the string.

checkPrefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

#renaming f2 as removeLastElement, since it removes the last element.
removeLastElement <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

#renaming function f3 replicateNTimes because it uses the rep function to replicate y the length of x times.
replicateNTimes <- function(x, y) {
  rep(y, length.out = length(x))
}
```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

```{r}
rnorm
MASS::mvrnorm

#rnorm has default values for mean and sd, not for n
#mvrnorm has default values for n, tol, empirical and EISPACK
#mvrnorn does not have default values for mu and sigma
# we could get rid of EISPACK from mvrnorm as it looks like it was used in an older version but no longer used.
# this would mean the only extra input between the two is tol and empirical
# mvrnorm has more stop statements that rnorm -- if checks for positive sigma, etc. were removed this would simplify the function and make it more similar to rnorm.
# it looks like the call in rnorm is to pass objects to C code. mvrnorm doesn't have a call to C


?rnorm
?mvrnorm
?.Call
```

3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

```{r}
#using lapply to apply a function that finds the coefficient of variation for all columns in dataset
lapply(mtcars, function(x) sd(x)/mean(x))

```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
    
```{r}
df <- data.frame(col1 = c(1, 2, 3, 4, 5, 6, 7, 8), col2 = c(5, 6, 2, 6, 4, 1, 3, 1), col3 = c(3, 4, 8, 5, 7, 1, 9, 10))

# SD of every column in the numeric data frame df
vapply(df, FUN=sd, numeric(1))

df2 <- data.frame(col1 = c(1, 2, 3, 4, 5, 6, 7, 8), col2 = c(5, 6, 2, 6, 4, 1, 3, 1), col3 = c(3, 4, 8, 5, 7, 1, 9, 10), col4 = c("a", "b", "c", "c", "b", "b","a", "b"))

# checks which columns are numeric for df2, and then finds sd for those columns of df2
v1 <- vapply(df2, is.numeric, logical(1))
vapply(df2[,v1], sd, numeric(1))

```
